---
title: "International Severe Acute Respiratory and Emerging Infections Consortium (ISARIC)"
output:
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float:
      collapsed: no
  pdf_document:
    toc: yes
    toc_depth: '2'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(tidyverse)
library(glue)
library(gridExtra)

# file locations

paths <- read_csv("paths.csv", col_names = F)

code.path <- paths[[which(paths$X1 == "code.path"), 2]]
data.path <- paths[[which(paths$X1 == "data.path"), 2]]
data.dict.file <- paths[[which(paths$X1 == "data.dict.file"), 2]]
site.list.file <- paths[[which(paths$X1 == "site.list.file"), 2]]
uk.data.file <- paths[[which(paths$X1 == "uk.data.file"), 2]]
row.data.file <- paths[[which(paths$X1 == "row.data.file"), 2]]
eot.data.file <- paths[[which(paths$X1 == "eot.data.file"), 2]]



# Source files

source(glue("{code.path}/shared.R"))
source(glue("{code.path}/output.R"))

```

<!-- <p style="font-family: times, serif; font-size:24pt> **International Severe Acute Respiratory and Emerging Infections Consortium (ISARIC)**</p> -->

<!-- <f style="font-family: times, serif; font-size:11pt; font-style:italic"> - A global federation of clinical research networks, providing a proficient, coordinated, and agile research response to outbreak-prone infectious diseases</f> -->


<f style="font-family: times, serif; font-size:11pt; font-style:italic"> - A global federation of clinical research networks, providing a proficient, coordinated, and agile research response to outbreak-prone infectious diseases</f>

# COVID-19 Report: 23rd March, 2020 {-}



# Summary

The results in this report have been produced using data from the International Severe Acute Respiratory and Emerging Infection Consortium (ISARIC) COVID-19 database up to 19 March 2020. These data were contributed by **`r de$N.sites`** sites across **`r de$N.countries`** countries.

To date, data have been entered for **`r de$N.cases`** patients. 

The cohort is made up of `r de$N.males` males and `r de$N.females` females - sex is unreported for `r de$N.sex.unknown` cases. 

The median age (calculated based on reported ages) is `r de$median.age` years. The minimum and maximum observed ages in this cohort are `r de$min.age`  and `r de$max.age` years respectively. 

Outcomes have been recorded for `r de$N.outcomes` patients, consisting of `r de$N.recoveries` recoveries and `r de$N.deaths` deaths. Follow-up is ongoing for `r de$N.censored` patients.

For this cohort,  the observed average for the number of days from admission to any outcome (death or discharge) is `r de$obs.mean.adm.outcome`, with a standard deviation of `r  de$obs.sd.adm.outcome`. 

The observed average for the number of days from (first) symptom onset to hospital admission is `r de$obs.mean.onset.to.adm`(95% CI: `r de$mean.onset.to.adm.lower`, `r de$mean.onset.to.adm.upper`) with a standard deviation of `r de$obs.sd.onset.adm`. The expected estimate for the mean for the same variable is `r de$mean.onset.to.adm` days (95% CI: `r de$mean.onset.to.adm.lower`, `r de$mean.onset.to.adm.upper`).

The average duration (in days) from admission to ICU is `r round(mean(de$a.icu, na.rm=T), 1)`, with a standard deviation of	`r round(sd(de$a.icu, na.rm=T), 1)`days. This is based on `r length(de$a.icu)` observations. The duration of ICU stay has an average of `r round(mean(de$d.icu, na.rm=T), 1)`and a standard deviation of	`r round(sd(de$d.icu, na.rm=T), 1)`.

The mean duration (in days) from admission to receiving NIMV is `r round(mean(de$a.nimv, na.rm=T), 1)` days, with a standard deviation of `r round(sd(de$a.nimv, na.rm=T), 1)` days.	This is	based on `r length(de$a.nimv)` observations.	The mean duration of NIMV is `r round(mean(de$d.nimv, na.rm=T), 1)` days, with a standard deviation (SD) of 	`r round(sd(de$d.nimv, na.rm=T), 1)` days. 


For the time from admission to IMV, the average duration (in days) is `r round(mean(de$a.imv, na.rm=T), 1)` (SD:`r round(sd(de$a.imv, na.rm=T), 1)`). The duration of IMV has an average of `r round(mean(de$d.imv, na.rm=T), 1)`	(SD:`r round(sd(de$d.imv, na.rm=T), 1)`). 	
	




Using only completed outcomes, the case fatality ratio (CFR) was estimated as `r de$hfr` (95% CI: `r de$hfr.lower`, `r de$hfr.upper`). When cases with unobserved outcomes are taken into consideration, the CFR estimate is  `r de$cfr` (95% CI: `r de$cfr.lower`, `r de$cfr.upper`).  

Of `r de$n.treat` patients with completed details of treatments received `r de$p.none`% of patients did not receive any antimicrobial treatments or steroids. `r de$p.abx`% received an antibiotic and `r de$p.av`% received antivirals.




# Patient Characteristics


**Figure 1**: Age and sex distribution of patients. Bar fills are outcome (death/discharge/censored) at the time of report.

```{r agepyramid, echo = FALSE}

age.pyramid(patient.data)

```

**Figure 2**: Top: The distribution of combinations of the four most common symptoms seen at admission, across all patients for whom this data was recorded. Filled and empty circles below the x-axis indiciate the presence or absence of each symptoms. Bottom: Frequency of each symptom in the patient population.

```{r symptomcombos, echo=FALSE, out.width= "100%", fig.height=3, message=FALSE}

symptoms.upset(patient.data, 4)
symptom.prevalence(patient.data)$plt

```


**Figure 3**:  Top: The distribution of combinations of the four most common comorbidities seen at admission, across all patients for whom this data was recorded. Filled and empty circles below the x-axis indiciate the presence or absence of each comorbidity. Bottom: Frequency of each comorbidity in the patient population.

```{r comorbcombos, echo=FALSE, out.width= "100%", fig.height=3, message=FALSE}

comorbidities.upset(patient.data, 4)
comorbidity.prevalence(patient.data)$plt

```

**Figure 4**: The distribution of patient status by number days after admission. "Transferred" patients were moved to another institution, where their final outcome cannot be determined. Patients with censored outcomes at the time of this report remain in the "Admitted" category. 

```{r outcome.curve, out.width= "80%"}

  
status.by.time.after.admission(patient.data)
```

**Figure 5**: Distribution of length of hospital stay, according to patient sex.

```{r stay.by.sex, out.width= "80%"}

violin.sex.func(patient.data)

```

**Figure 6**: Distribution of length of hospital stay, according to patient age group. 

```{r  stay.by.age, out.width= "80%"}

violin.age.func(patient.data)
```


<!-- **Figure 7**: Estimated hospital fatality ratio based on patients with complete outcome data, showing changes to estimate over time. Shaded area represents 95% confidence interval. Methods from Wu *et al.*(2020). -->

<!-- ```{r out.width= "80%"} -->

<!-- hospital.fatality.ratio(patient.data)$plt -->

<!-- ``` -->

**Figure 7**: Patient outcomes by epidemiological week (of 2020) of admission

```{r recruitment3, echo=FALSE}

outcomes.by.admission.date(patient.data)

```


# Treatment

**Figure 8**: Treatments used on all patients where this information was recorded.

```{r treatment1, echo=FALSE}
treatment.use.plot(patient.data)$plt
```


**Figure 9**: The distribution of combinations of antimicrobial treatments and steroids administed during hospital stay, across all patients for whom this data was recorded. Filled and empty circles below the x-axis indiciate the presence or absence of each symptoms.

```{r treatment2, echo=FALSE}
treatment.upset(patient.data)$p
```


# Statistical Analysis

Blue curves show the Gamma distribution fit to two time-dependent data variables. 


**Figure 10**: Distribution of time from symptom onset to admission.

```{r onset.admission, out.width = '80%'}

onset.adm.func(patient.data)$plt

```


**Figure 11**: Distribution of time from admission to an outcome - either death or recovery.


```{r  adm.outcome, out.width = '80%'}

adm.outcome.func(patient.data)$plt

```



**Figure 12**: Nonparametric probabilites for death (red curve) and  recovery (blue curve) over time. The case  fatality ratio (black) is based on the cumulative incidence function of deaths and recoveries so far. See Methods section for details. For a completed epidemic, the curves for death and recovery meet. Estimates were derived using a nonparametric Kaplan-Meier–based method proposed by Ghani *et al.* (2005).


```{r outcomelines}

modified.km.plot(patient.data)

```


# Country Comparisons

**Figure 13**: Number of sites per country

```{r recruitment1, echo=FALSE}

sites.by.country(patient.data)

```

**Figure 14**: Distribution of patients by country and outcome

```{r recruitment2, echo=FALSE}

outcomes.by.country(patient.data)

```




# Background


In response to the emergence of novel coronavirus (COVID-19), ISARIC launched a portfolio of resources to accelerate outbreak research and response. All resources are designed to address the most critical public health questions, have undergone extensive review by international clinical experts, and are free for all to use. Resources are available on the [ISARIC website](https://isaric.tghn.org/).

The [ISARIC-WHO COVID-19 Case Record Form (CRF)](https://media.tghn.org/medialibrary/2020/03/ISARIC_COVID-19_CRF_V1.3_24Feb2020.pdf) enables the collection of standardised clinical data to inform patient management and the public health response. These forms should be used to collect data on suspected or confirmed cases of COVID-19.

Following the launch of this this open resource, ISARIC received a massive response from the health and research communities. This form is now in use across dozens of countries and research consortia, aligning the collection of health data. The form is available in multiple languages. Data can be entered electronically to the data platform as below. ISARIC supports researchers to retain control of the data and samples they collect. All decisions regarding data use are made by the institutions that enter the data. We keep our contributors informed of any plans and welcome their input to ensure that we are generating the best science and promoting the interests of your patients, your institutions and your public health authorities. Feedback and suggestions are welcome anytime at [ncov@isaric.org](ncov@isaric.org)

To support the rapid implementation of standardised data collection and reporting, ISARIC hosts a data platform that includes an electronic data capture system, a secure repository and an analytic framework. Data are entered to a web-based [REDCap](https://www.project-redcap.org/) data management system, securely stored, and used to inform regular reports as above. Data contributors are invited to input on the methods and contents of the reports, and are provided with the R code to execute analysis on their own data in the platform. For more information, visit the [ISARIC](https://isaric.tghn.org/covid-19-clinical-research-resources/) website.




# Methods

Patient details were submitted electronically by participating sites to the ISARIC database.  Relevant background and presenting symptoms were recorded on the day of study recruitment.  Daily follow-up was then completed until recovery or death.  A final form was completed with details of treatments received and outcomes. All categories that represent fewer than five individuals have been suppressed to avoid the potential for identification of participants. 

Graphs have been used to represent the age distribution of patience by sex and status (censored, death, recovered), the prevalence of individual symptoms  - and combinations of them - on admission, the prevalence of individual comorbidities  - and combinations of them - on admission, the length of hospital stay by sex and age group and the distribution of patient statuses by time since admission. In addition, the number of cases recruited by country and site, as well as the case count by status, has been represented.

Using a non-parametric Kaplan-Meier-based method (Ghani *et al.*, 2005), the case- fatality ratio (CFR) was estimated, as well as probabilities for death and recovery. This method estimates the CFR with the formula $a/(a+b)$, where $a$ and $b$ are the values of the cumulative incidence function for deaths and recoveries respectively, estimated at the last observed time point. In a competing risk context (i.e. where there are multiple endpoints),  the cumulative incidence function for an endpoint is equal to the product of the hazard function for that endpoint and the survival function assuming a composite endpoint. It is worth noting that this method assumes that future deaths and recoveries will occur with the same relative probabilities as have been observed so far. Binomial confidence intervals for the CFR were obtained by a normal approximation (See Ghani *et al.*, (2005)).

<!-- Another CFR estimation method, which uses observed outcomes only, was employed in this analysis. For this method, the CFR is calculated as the ratio of deaths to the sum of deaths and recoveries until the reference point (Wu *et al.*, 2020). Over time, the uncertainty around the estimate decreases due to an increase in completed outcomes. Exact binomial confidence intervals are plotted along with the risk estimate at each time point.  -->

To obtain estimates for the distributions of time from symptom onset to hospital admission and the time from admission to outcome (death or recovery), Gamma distributions were fitted to the observed data, accounting for censorship. Parameters were estimated by a maximum likelihood procedure and confidence intervals for the means and variances were obtained by bootstrap. 

<!-- For cases still in hospital, i.e. cases which do not yet have an outcome by the date of this report, the  likelihood function for the estimation of the time from admission to any outcome was constructed to allow for right censoring. -->

All analysis were performed using the R statistical software (R Core Team, 2019).



# Caveats

## Caveat to all data

Patient data is collected and uploaded from start of admission, however a complete patient data set is not available until the episode of care is complete. This causes a predictable lag in available data influenced by the duration of admission which is greatest for the sickest patients, and accentuated during the up-phase of the outbreak.


## Caveat to UK data

In the UK, up to 10th March 2020, people with positive swabs were admitted to hospital as part of the containment strategy. Many of these people had recently been abroad, particularly Italy. Since 11th March 2020, admission is mostly based upon need for treatment of COVID-19 disease. This may explain the younger age of the UK group and the fewer comorbidities reported when compared with reports from China and Italy.



# Summary Tables


**Table 1:** Patient Characteristics

| Description                  | Value                                      |
|------------------------------|--------------------------------------------|
| Size of cohort               | `r de$N.cases`                            |
|                              |                                            |
| **By sex**                      |                                            |
| Male                         | `r de$N.males`                            |
| Female                       | `r de$N.females`                          |
| Unknown                      | `r de$N.sex.unknown`                     |
|                              |                                            |
| **By Status**                   |                                            |
| Dead                         | `r de$N.deaths` (`r round(de$N.deaths/de$N.cases,2)`) |
| Recovered (discharged alive) | `r de$N.recoveries`                          |
| Still in hospital            | `r de$N.censored`                            |
|                              |                                            |
| **By age group**                 |                                            |
| `r levels(patient.data$agegp10)[1]` | `r summary(patient.data$agegp10)[[1]]`          |
| `r levels(patient.data$agegp10)[2]` | `r summary(patient.data$agegp10)[[2]]`             |
| `r levels(patient.data$agegp10)[3]` | `r summary(patient.data$agegp10)[[3]]`             |
| `r levels(patient.data$agegp10)[4]` | `r summary(patient.data$agegp10)[[4]]`             |
| `r levels(patient.data$agegp10)[5]` | `r summary(patient.data$agegp10)[[5]]`             |
| `r levels(patient.data$agegp10)[6]` | `r summary(patient.data$agegp10)[[6]]`             |
| `r levels(patient.data$agegp10)[7]` | `r summary(patient.data$agegp10)[[7]]`             |
| `r levels(patient.data$agegp10)[8]` | `r summary(patient.data$agegp10)[[8]]`             |
|                              |                                                      |
| Unknown                      | `r summary(patient.data$agegp10)[[9]]`             |
|                              |                                            |
| **By ethnicity**                 |                                            |
| Arab                         | `r de$eth.1`                                  |
| Black                        | `r de$eth.2`                                  |
| East Asian                   | `r de$eth.3`                                  |
| South Asian                  | `r de$eth.4`                                  |
| West Asian                   | `r de$eth.5`                                  |
| Latin American               | `r de$eth.6`                                  |
| White                        | `r de$eth.7`                                  |
| Aboriginal/First Nations     | `r de$eth.8`                                  |
| Other                        | `r de$eth.9`                                  |
| Unknown                      | `r de$eth.NA`                                 |


<br /><br /><br />


**Table 2:** Prevalence of Symptoms, Comorbidities and Treatments

| **Symptoms**                      | Present                        | Absent                        | Unknown                        |
|-------------------------------|--------------------------------|-------------------------------|--------------------------------|
| Abdominal pain                | `r de$s.dat$present[1]`        | `r de$s.dat$absent[1]`        | `r de$s.dat$unknown[1]`        |
| Bleeding                      | `r de$s.dat$present[2]`        | `r de$s.dat$absent[2]`        | `r de$s.dat$unknown[2]`        |
| Chest pain                    | `r de$s.dat$present[3]`        | `r de$s.dat$absent[3]`        | `r de$s.dat$unknown[3]`        |
| Confusion                     | `r de$s.dat$present[4]`        | `r de$s.dat$absent[4]`        | `r de$s.dat$unknown[4]`        |
| Conjuctivitis                 | `r de$s.dat$present[5]`        | `r de$s.dat$absent[5]`        | `r de$s.dat$unknown[5]`        |
| Cough                         | `r sum(de$s.dat$present[6:8])` | `r sum(de$s.dat$absent[6:8])` | `r sum(de$s.dat$unknown[6:8])` |
| Diarrhoea                     | `r de$s.dat$present[9]`        | `r de$s.dat$absent[9]`        | `r de$s.dat$unknown[9]`        |
| Ear pain                      | `r de$s.dat$present[10]`       | `r de$s.dat$absent[10]`       | `r de$s.dat$unknown[10]`       |
| Fatigue                       | `r de$s.dat$present[11]`       | `r de$s.dat$absent[11]`       | `r de$s.dat$unknown[11]`       |
| Fever                         | `r de$s.dat$present[12]`       | `r de$s.dat$absent[12]`       | `r de$s.dat$unknown[12]`       |
| Headache                      | `r de$s.dat$present[13]`       | `r de$s.dat$absent[13]`       | `r de$s.dat$unknown[13]`       |
| Joint pain                    | `r de$s.dat$present[14]`       | `r de$s.dat$absent[14]`       | `r de$s.dat$unknown[14]`       |
| Lymph                         | `r de$s.dat$present[15]`       | `r de$s.dat$absent[15]`       | `r de$s.dat$unknown[15]`       |
| Myalgia                       | `r de$s.dat$present[16]`       | `r de$s.dat$absent[16]`       | `r de$s.dat$unknown[16]`       |
| Rash                          | `r de$s.dat$present[17]`       | `r de$s.dat$absent[17]`       | `r de$s.dat$unknown[17]`       |
| Runny nose                    | `r de$s.dat$present[18]`       | `r de$s.dat$absent[18]`       | `r de$s.dat$unknown[18]`       |
| Seizures                      | `r de$s.dat$present[19]`       | `r de$s.dat$absent[19]`       | `r de$s.dat$unknown[19]`       |
| Shortness of breath           | `r de$s.dat$present[20]`       | `r de$s.dat$absent[20]`       | `r de$s.dat$unknown[20]`       |
|                               |                                |                               |                                |
| **Comorbidities**                 |                                |                               |                                |
| AIDS/HIV                      | `r de$c.dat$present[1]`        | `r de$c.dat$absent[1]`        | `r de$c.dat$unknown[1]`        |
| Asthma                        | `r de$c.dat$present[2]`        | `r de$c.dat$absent[2]`        | `r de$c.dat$unknown[2]`        |
| Chronic cardiac disease       | `r de$c.dat$present[3]`        | `r de$c.dat$absent[3]`        | `r de$c.dat$unknown[3]`        |
| Chronic haematologic disease  | `r de$c.dat$present[4]`        | `r de$c.dat$absent[4]`        | `r de$c.dat$unknown[4]`        |
| Chronic neurological disorder | `r de$c.dat$present[5]`        | `r de$c.dat$absent[5]`        | `r de$c.dat$unknown[5]`        |
| Chronic pulmonary disease     | `r de$c.dat$present[6]`        | `r de$c.dat$absent[6]`        | `r de$c.dat$unknown[6]`        |
| Dementia                      | `r de$c.dat$present[7]`        | `r de$c.dat$absent[7]`        | `r de$c.dat$unknown[7]`        |
| Diabetes                      | `r de$c.dat$present[8]`        | `r de$c.dat$absent[8]`        | `r de$c.dat$unknown[8]`        |
| Liver disease                 | `r de$c.dat$present[9]`        | `r de$c.dat$absent[9]`        | `r de$c.dat$unknown[9]`        |
| Malignant neoplasm            | `r de$c.dat$present[10]`       | `r de$c.dat$absent[10]`       | `r de$c.dat$unknown[10]`       |
| Malnutrition                  | `r de$c.dat$absent[11]`        | `r de$c.dat$absent[11]`       | `r de$c.dat$unknown[11]`       |
| Obesity                       | `r de$c.dat$absent[12]`        | `r de$c.dat$absent[12]`       | `r de$c.dat$unknown[12]`       |
| Chronic kidney disease        | `r de$c.dat$present[14]`       | `r de$c.dat$absent[14]`       | `r de$c.dat$unknown[14]`       |
| Rheumatologic disorder        | `r de$c.dat$present[15]`       | `r de$c.dat$absent[15]`       | `r de$c.dat$unknown[15]`       |
| Smoking                       | `r de$c.dat$present[16]`       | `r de$c.dat$absent[16]`       | `r de$c.dat$unknown[16]`       |
| Other risk factors            | `r de$c.dat$present[13]`       | `r de$c.dat$absent[13]`       | `r de$c.dat$unknown[13]`       |
|                               |                                |                               |                                |
| **Treatment**                     |                                |                               |                                |
| Antibiotic agent              | `r de$t.dat$present[1]`        | `r de$t.dat$absent[1]`        | `r de$t.dat$unknown[1]`        |
| Antifungal agent              | `r de$t.dat$present[2]`        | `r de$t.dat$absent[2]`        | `r de$t.dat$unknown[2]`        |
| Antiviral agent               | `r de$t.dat$present[3]`        | `r de$t.dat$absent[3]`        | `r de$t.dat$unknown[3]`        |
| Corticosteroid agent          | `r de$t.dat$present[4]`        | `r de$t.dat$absent[4]`        | `r de$t.dat$unknown[4]`        |
| Extracorporeal support        | `r de$t.dat$present[5]`        | `r de$t.dat$absent[5]`        | `r de$t.dat$unknown[5]`        |
| Inhaled nitric oxide          | `r de$t.dat$present[6]`        | `r de$t.dat$absent[6]`        | `r de$t.dat$unknown[6]`        |
| Inotropes / vasopressors      | `r de$t.dat$present[7]`        | `r de$t.dat$absent[7]`        | `r de$t.dat$unknown[7]`        |
| Invasive ventilation          | `r de$t.dat$present[8]`        | `r de$t.dat$absent[8]`        | `r de$t.dat$unknown[8]`        |
| Non-invasive ventilation      | `r de$t.dat$present[9]`        | `r de$t.dat$absent[9]`        | `r de$t.dat$unknown[9]`        |
| Oxygen therapy                | `r de$t.dat$present[11]`       | `r de$t.dat$absent[11]`       | `r de$t.dat$unknown[11]`       |
| Prone ventilation             | `r de$t.dat$present[12]`       | `r de$t.dat$absent[12]`       | `r de$t.dat$unknown[12]`       |
| Renal replacement therapy     | `r de$t.dat$present[13]`       | `r de$t.dat$absent[13]`       | `r de$t.dat$unknown[13]`       |
| Tracheostomy inserted         | `r de$t.dat$present[14]`       | `r de$t.dat$absent[14]`       | `r de$t.dat$unknown[14]`       |
| Other                         | `r de$t.dat$present[10]`       | `r de$t.dat$absent[10]`       | `r de$t.dat$unknown[10]`       |

<br /><br /><br />


**Table 3:** Duration of stay


| **Time(in days)**   | **Observed mean**  | **Observed SD**    | **Expected mean**^[Expected estimates were derived by fitting a Gamma distribution to the observed data.] (95% CI)** | **Notes** |
|------------------------------------------------|----------------------------------------------------------------------------------------------------|--------------------------------------|-------------------------------------------------------------------------------------------|---------------------------------------------|
| Length of hospital stay              | `r de$obs.mean.adm.outcome` (`r de$obs.mean.adm.outcome.lower`, `r de$obs.mean.adm.outcome.upper`) | `r de$obs.sd.adm.outcome`            | `r de$mean.adm.to.outcome` (`r de$adm.outcome.lower`, `r de$adm.outcome.upper`)           |                                             |
|Symptom onset to admission | `r  de$obs.mean.onset.adm` (`r de$obs.mean.onset.adm.lower`, `r de$obs.mean.onset.adm.upper`)      | `r de$obs.sd.onset.adm`              | `r de$mean.onset.to.adm` (`r de$mean.onset.to.adm.lower`, `r de$mean.onset.to.adm.upper`) |                                             |
|                                                |                                                                                                    |                                      |                                                                                           |                                             |
|Admission to ICU entry     | `r round(mean(de$a.icu, na.rm=T), 1)`                                                              | `r round(sd(de$a.icu, na.rm=T), 1)`  |                                                                                           | Based on `r length(de$a.icu)` observations  |
| Duration of ICU                     | `r round(mean(de$d.icu, na.rm=T), 1)`                                                              | `r round(sd(de$d.icu, na.rm=T), 1)`     |                                                                                           | Based on `r length(de$d.icu)` observations  |
|                                                |                                                                                                    |                                      |                                                                                           |                                             |
| Admission to IMV           | `r round(mean(de$a.imv, na.rm=T), 1)`                                                              | `r round(sd(de$a.imv, na.rm=T), 1)`  |                                                                                           | Based on `r length(de$a.imv)` observations  |
| Duration of IMV                                | `r round(mean(de$d.imv, na.rm=T), 1)`                                                              | `r round(sd(de$d.imv, na.rm=T), 1)`  |                                                                                           | Based on `r length(de$d.imv)` observations  |
|                                                |                                                                                                    |                                      |                                                                                           |                                             |
| Admission to NIMV          | `r round(mean(de$a.nimv, na.rm=T), 1)`                                                             | `r round(sd(de$a.nimv, na.rm=T), 1)` |                                                                                           | Based on `r length(de$a.nimv)` observations |
| Duration of NIMV                               | `r round(mean(de$d.nimv, na.rm=T), 1)`                                                             | `r round(sd(de$d.nimv, na.rm=T), 1)` |                                                                                           | Based on `r length(de$d.nimv)` observations |

The expected mean for the same variable is `r de$mean.adm.to.outcome` days (95% CI: `r de$adm.outcome.lower`, `r de$adm.outcome.upper`).

# ISARIC Team Members

* Gail Carson
* Emmanuelle A. Dankwa
* AnneMarie Docherty 
* Christl A. Donnelly
* Christophe Fraser
* Matthew Hall
* Peter Horby
* Kalynn Kennon
* James Lee
* Laura Merson
* Sarah Moore
* Piero Olliaro
* Daniel Plotkin
* Mark Pritchard
* Calum Semple 





# References

1.	A. C. Ghani, C. A. Donnelly, D. R. Cox, J. T. Griffin, C. Fraser, T. H. Lam, L. M. Ho, W. S. Chan, R. M. Anderson, A. J. Hedley, G. M. Leung (2005).  Methods for Estimating the Case Fatality Ratio for a Novel, Emerging Infectious Disease, *American Journal of Epidemiology*, 162(5), 479 - 486. [doi:10.1093/aje/kwi230](https://doi.org/10.1093/aje/kwi230).

2.	R Core Team (2019). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria.

3.	Wu Peng, Hao Xinxin, Lau Eric H Y, Wong Jessica Y, Leung Kathy S M, Wu Joseph T, Cowling Benjamin J, Leung Gabriel M. Real-time tentative assessment of the epidemiological characteristics of novel coronavirus infections in Wuhan, China, as at 22 January 2020 (2020). *Euro Surveill.*, 25(3):pii=2000044. [doi:10.2807/1560-7917.ES.2020.25.3.2000044](https://doi.org/10.2807/1560-7917.ES.2020.25.3.2000044).






